\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{braket}
% \usepackage[utf8]{inputenc}
\newcommand{\angstrom}{\mbox{\normalfont\AA}}
\begin{document}

\title{Discrete Variable Representation Introduction and Derivation}
\author{Luke Hatcher}
\date{University of Washington: Department of Chemistry \\ McCoy Group, Winter 2020}
\maketitle


\section{Discrete Variable Representation}
\subsection{Introduction}
Discrete Variable Representation (DVR) solves the time independent Schrodinger equation.  
\subsection{Basis functions}
For DVR, a finite set of basis functions must be chosen. 
For this case our set $\{\phi_i\}_N$ belongs to a space such that they satisfy
\begin{equation}
\braket{\phi_i|\phi_j} = \delta_{ij}
\end{equation}
and thus are orthonormal.  
It follows that we can represent our wavefunction as
\begin{equation}
\psi=\sum_i{c_i}\phi_i
\end{equation}

\begin{equation}
c_i = \braket{\phi_i|\psi}
\end{equation}

\subsection{Matrix Representations}
Starting from the time independent Schrodinger equation
\begin{equation}
\hat{H}\psi = E\psi
\end{equation}
recall that
\begin{equation}
\hat{H}=\hat{T}+\hat{V}
\end{equation}
We want to represent our Hamiltonian as the sum of two matrices.
Thanks to Colbert and Miller, the matrix elements of the kinetic energy matrix are analytically known.$^{[1]}$
\begin{equation}
\hat{\textbf{T}}_{ij} = \frac{\hbar^2(-1)^{i-j}}{2m\Delta{x^2}}
\begin{cases}
\frac{\pi^2}{3}, &i=j\\
\frac{2}{(i-j)^2}, &i\neq{j}
\end{cases}
\end{equation}
It follows that the potential energy matrix is diagonal. 
\begin{equation}
\hat{\textbf{V}}_{ij}=\delta_{ij}V(x_i)
\end{equation}
Here the grid points $x_i$ are evenly spaced by $\Delta{x}$ and the size of the grid is appropriately large so that the wavefunction tends to zero at both ends.
Doing so will represent an interval of $(-\infty, \infty)$.
As a reminder, the number of grid points is equal to the Gaussian 09 input for scan points along the bond of interest. 
For example 120 points along the shared OH as mentioned in section 1.3.
Likewise, the potential at each point $V(x_i)$ is provided by Gaussian 09.
Now, we can add our kinetic and potential matrix representations together to get our Hamiltonian matrix.
\begin{equation}
\hat{\textbf{H}}=\hat{\textbf{T}}+\hat{\textbf{V}}
\end{equation}

\subsection{Solving for $\psi$}
Now all that remains is an eigenvalue problem.
\begin{equation}
\hat{\textbf{H}}\psi=\epsilon_n\psi
\end{equation}

\begin{equation}
(\hat{\textbf{H}}-\epsilon_\lambda\textbf{I})\psi=0
\end{equation}
Nontrivial solutions exist when 
\begin{equation}
det(\hat{\textbf{H}}-\epsilon_\lambda\textbf{I})=0
\end{equation}
which after solving provides us with our desired energy eigenvalues $\epsilon_n$ and thus their respected eigenvectors $\psi$.

\section{References}
[1] D. T. Colbert and W. H. Miller, J. Chem. Phys. 96, 1982 (1992).




\end{document}



